EuPPollNet: A European database of plant-pollinator networks

### 2024-09-16 Manuscript/EuPPollNet_Ms.qmd
edits

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022. Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019]. This approach maximized the potential number of studies that could be incorporated in this database. The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*". To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though the pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).

@@

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations). The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet. Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%). All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans. The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions. Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). Species that belong to other orders `r format(round(100-spp_percent_main_order, 2), nsmall = 2)`% are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%). Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

@@

The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble) + 1` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species. These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29). The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30). Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies. To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year. Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.

@@

@@

To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network in binary form (i.e., presence-absence) and examined how these network metrics change across different latitudes and bioclimatic regions in Europe. We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation. We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics. As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance. The relationship between residual connectance and species richness was investigated using beta regression. This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1. The model was implemented using the betareg package [@cribari2010]. We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017]. This metric was calculated using the **maxnodf** package [@hoeppke2021]. Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude. In addition, to quantify how connectance and nestedness change with network size, we determined their association with the number of species per network using the Kendall rank correlation coefficient.

Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008. The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions. We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008]. These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals. Both connectance and nestedness (NODF) were estimated for each network by using the function _networklevel_ from **bipartite**.

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}. Phylogenetic and taxonomic coverage of the plant families at European level. The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale. Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization. The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level. Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}. Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their unique pairwise interactions across networks. Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks. The solid black points indicate the number of species and interactions contained in the database. Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species. This last graph uses the same color and shape structure as the ones in the top panel. Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks. Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network. Note that \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}

@@

#Fix levels
d_int1_lines = d_int1_lines %>% 
mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
                       Method == "Extrapolation" ~ "Extrapolated",
                       T ~ col))
#Order levels
d_int1_lines$col = factor(d_int1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
#Plot
p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = Method, color = col)) +
geom_line(aes(linetype = col), size = 0.65) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_int1_point, col = "black", size=2) + 
ylab("Unique interactions") +
xlab("Networks") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off")  +
xlim(0,3000) +
ylim(0, 30000)+
ggtitle("(c)")

@@

#Plot
o1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.03, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
xlab("Pollinator species") +
ylab("Occurrence across networks") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.35) + 
annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=200, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3)  +
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
                ylim=c(0,0.7), clip="off") +
ggtitle("(e)")

@@

o2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.03, fill = "#009E73") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
xlab("Plant species") +
ylab("Occurrence across networks") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.35) + 
annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=150, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3) + 
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
                ylim=c(0, 0.7), clip="off") +
ggtitle("(f)")

@@

#Fix levels
d_pp1_lines = d_pp1_lines %>% 
mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
                       Method == "Extrapolation" ~ "Extrapolated",
                       T ~ col))
#Order levels
d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
#Plot
pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = Method, color = col)) +
geom_line(aes(linetype = col)) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_pp1_point, col = "black") + 
ylab("Pollinator species") +
xlab("Plant species") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off") +
ggtitle("(d)") +
scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
scale_y_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2600)) +
theme(legend.position = "none")

@@

#Outputs calculated in script 7_3
#Load lm output connectance ~ latitude
connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
#Load lm output nestedness ~ latitude
normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
#Load values of connectance
min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
#Load values of nestedness
min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
#Load kendall corr from connectance and spp
connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
#Load kendall corr from nestedness and spp
normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
#Load z values
#Save data
z_percent = readRDS("../Data/Manuscript_info/z_percent.rds")
no_diff_z_percent = z_percent %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent = z_percent %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent = z_percent %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)
#Load correlation outputs of residual connectance and nestednes
resid_connectance_nestedness_corr_tau = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_tau.rds")
resid_connectance_nestedness_corr_pval = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_pval.rds")

@@

#Prepare now nestedness plots
# Plot
p = 0.05 #cutoff probability 95% confidence
critical_value = qnorm(p/2) #double tail probability divide by 2

>>>>

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022. Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019]. This approach maximized the potential number of studies that could be incorporated in this database. The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'. To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though the pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).

@@

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations). The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet. Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%). All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans. The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions. Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%). Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

@@

The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble)` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species. These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29). The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30). Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies. To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year. Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.

@@

#Small edit to column number
my_tibble = my_tibble %>% 
mutate(Variable = paste0(seq(1:length(Variable)),". ", Variable))

@@

To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network and examined how these network metrics change across different latitudes and bioclimatic regions in Europe. We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation. We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics. As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance. The relationship between residual connectance and species richness was investigated using a beta regression. This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1. The model was implemented using the betareg package [@cribari2010]. We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species in comparison to other nestedness metrics that change with network size [@song2017]. This metric was calculated using the **maxnodf** package [@hoeppke2021]. Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude. In addition, to quantify how connectance and nestedness change with network size, we determined their association with the number of species per network using the Kendall rank correlation coefficient.

Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008. The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions, avoiding the influence of network size. As NODF is a metric computed from binary matrices, we calculated 100 null models for the binarized version of each network using the 'curveball' algorithm [@strona2014]. This algorithm implements the configuration model (i.e. random rewiring of all links, without self-links or double links) and thus keeps the exact number of connections per species (i.e., realised degree). Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species. Each empirical network was randomized 100 times using each method. These null models were implemented with the **vegan** [@oksanen2013] package. Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite**.

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}. Phylogenetic and taxonomic coverage of the plant families at European level. The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale. Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization. The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level. Dark grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}. Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and the number of unique pairwise interactions across networks. Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks. The solid black points indicate the number of species and interactions contained in the database. Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species. This last graph uses the same color and shape structure as the ones in the top panel. Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks. Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network. Note that \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}

@@

#Fix levels
d_int1_lines = d_int1_lines %>% 
mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
                       Method == "Extrapolation" ~ "Extrapolated",
                       T ~ col))
#Order levels
d_int1_lines$col = factor(d_int1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
#Plot
p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = Method, color = col)) +
geom_line(aes(linetype = col), size = 0.65) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_int1_point, col = "black", size=2) + 
ylab("Unique interactions") +
xlab("Networks") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off")  +
xlim(0,3000) +
ylim(0, 35000)+
ggtitle("(c)")

@@

#Plot
o1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.03, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
xlab("Pollinator species") +
ylab("Occurrence across networks") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.35) + 
annotate(geom="text", x=1590, y=0.60, label="50th percentile",color="black", size = 3) +
annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=200, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3)  +
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
                ylim=c(0,0.7), clip="off") +
ggtitle("(e)")

@@

o2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.03, fill = "#009E73") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
xlab("Plant species") +
ylab("Occurrence across networks") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.35) + 
annotate(geom="text", x=1010, y=0.60, label="50th percentile",color="black", size = 3) + 
annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=150, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3) + 
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
                ylim=c(0, 0.7), clip="off") +
ggtitle("(f)")

@@

#Fix levels
d_pp1_lines = d_pp1_lines %>% 
mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
                       Method == "Extrapolation" ~ "Extrapolated",
                       T ~ col))
#Order levels
d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
#Plot
pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = Method, color = col)) +
geom_line(aes(linetype = col)) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_pp1_point, col = "black") + 
ylab("Pollinator species") +
xlab("Plant species") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off") +
ggtitle("(d)") +
scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
scale_y_continuous(breaks = seq(0, 3000, by = 1000), limits=c(0,3000)) +
theme(legend.position = "none")

@@

#Outputs calculated in script 7_3
#Load lm output connectance ~ latitude
connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
#Load lm output nestedness ~ latitude
normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
#Load values of connectance
min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
#Load values of nestedness
min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
#Load kendall corr from connectance and spp
connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
#Load kendall corr from nestedness and spp
normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
#Load z values
#VAZNULL
z_percent_vaznull = readRDS("../Data/Manuscript_info/z_percent_vaznull.rds")
no_diff_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)
#CURVEBALL
z_percent_curveball = readRDS("../Data/Manuscript_info/z_percent_curveball.rds")
no_diff_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)

#QUASISWAP
z_percent_quasiswap = readRDS("../Data/Manuscript_info/z_percent_quasiswap.rds")

no_diff_z_percent_quasiswap = z_percent_quasiswap %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_quasiswap = z_percent_quasiswap %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_quasiswap = z_percent_quasiswap %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)
#rd2
z_percent_rd2 = readRDS("../Data/Manuscript_info/z_percent_rd2.rds")

no_diff_z_percent_rd2 = z_percent_rd2 %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_rd2 = z_percent_rd2 %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_rd2 = z_percent_rd2 %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)

#Load correlation outputs of residual connectance and nestednes
resid_connectance_nestedness_corr_tau = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_tau.rds")
resid_connectance_nestedness_corr_pval = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_pval.rds")

@@

null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")

#Finally, find % of networks that differ statistically
# Merge and process data
d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
